Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2015-05-18 sas2015/content/impl.tex
review comments

%Let $P=(G^\#)$ be an IFDS problem and $Q=(G^\#,\,\edgefn)$ an IDE problem obtained by a conversion from $P$.
Let $G^\#$ be the exploded supergraph of an arbitrary IFDS problem. A \emph{transformation}
$\mathcal T:\ (G^\#)\to (G^\#,\,L,\,\edgefn)$
converts the IFDS problem into an IDE problem.
We consider two IFDS-to-IDE transformations: an \textit{equivalence transformation} $\transEq$ (pronounced ``t-equiv'') and
a \textit{correlated-calls transformation} $\transCC$ (pronounced ``t-c-c'').
Both transformations keep the exploded supergraph $G^\#$ the same, and only generate different edge functions.
The solution of the IDE problem
can be mapped back to an IFDS solution.
If the equivalence transformation was used, then this solution is identical to the
solution that would be computed by the IFDS algorithm for the original IFDS problem.
If the correlated-calls transformation was used, then this solution is more precise
because it excludes flow along infeasible paths due to correlated calls.

@@

%\paragraph{Lattice Elements}
In the correlated-calls transformation, the 
%Just like in the equivalence transformation $\transEq$, the exploded supergraph for $\transCC$ is the same as in the original IFDS problem. The 
lattice elements are maps from receivers to sets of types:
    $\lcc=\left\{\,m:\ R\to2^T\right\}$,
where $R$ is the set of receivers and $T$ is the set of all types.
For each receiver~$r$, the map gives an overapproximation of the possible
runtime types of $r$.
Sets of types are ordered by the superset relation, and this is lifted to maps from receivers
to sets of types, so the bottom element $\botcc$ maps every receiver to the set of all types,
and the top element $\topcc$ maps every receiver to the empty set of types.
During an actual execution, every receiver $r$ points to an object of some
runtime type. Therefore, a data-flow fact
is unreachable along a given path if its corresponding lattice element
maps any receiver to the empty set of types.

@@

A micro-function $f\in\lcc\to \lcc$ defines how the map from receivers to types should
be updated when an instruction is executed.
The micro-function for most kinds of instructions is the identity.
On a call to and return from a specific method $m$ called on receiver $r$,
the micro-function restricts the receiver-to-type map to map $r$ only to
types consistent with the polymorphic dispatch to method $m$.
Finally, when an instruction assigns an object of unknown type to a receiver
$r$, the corresponding micro-function updates the map to map $r$ to the
set of all types. This is made precise by the following definition:
\begin{definition}\label{def:microfn}
Given a previously fixed set $S \subseteq R$ of receivers, the micro-function $\varepsilon_S(e)$
of a supergraph edge $e$ is defined as:
                    %\otodo{define $n_1(e)$ globally somewhere early on, and consider using a more decriptive name like src instead of $n_1$}
  \begin{align}\label{eq:varepsilon}
      &\varepsilon_S(e)= \lambda m\,.\,\\
        &\begin{cases}
            m[r\to m(r)\cap \tau(s,\,f)],
            &\text{\parbox{75mm}{
                    if $e$ is a call-start edge $r.c() \to \textsf{start}_f$ that calls
                procedure $f$ with signature $s$,
                and $r\in S$;}}\\[3ex]
        \parbox{30mm}{\[\begin{array}{l}
                m[r\to m(r)\cap \tau(s,\,f)]\\
            [v_1\to\bot_T] \ldots [v_k\to\bot_T],
        \end{array}
        \]
        }
        &
        \text{\parbox{75mm}{
                if $e$ is an end-return edge
                $\textsf{end}_f \to \textsf{return}_{r.c()}$
                from method $f$ with signature $s$ to the return node corresponding
                to the call $r.c()$,
                $v_1,\,\dots,\,v_k\in S$ are the local variables
                in $f$,
                and $r\in S$;}}\\
            m\left[r\to \bot_T\right],
            &\text{\parbox{75mm}{if $e = n_1 \to n_2$ and $n_1$ contains an assignment
                    to $r\in S$;}}\\
            m
                &\text{otherwise.}
        \end{cases}
  \end{align}
\end{definition}

@@

\begin{restatable}[Soundness]{lemma}{sound}\label{lem:sound}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  $
    d\in \backCC\left(\resultIDE(\transCC_R(P))\right)(n)\,.
  $
\end{restatable}

>>>>

%Let $P=(G^\#)$ be an IFDS problem and $Q=(G^\#,\,\edgefn)$ an IDE problem obtained by a conversion from $P$.
Let $G^\#$ be the exploded supergraph of an arbitrary IFDS problem. A \emph{transformation}
$\mathcal T:\ (G^\#)\to (G^\#,\,L,\,\edgefn)$
converts the IFDS problem into an IDE problem.
We consider two IFDS-to-IDE transformations: an \textit{equivalence transformation} $\transEq$ (pronounced ``t-equiv'') and
a \textit{correlated-calls transformation} $\transCC_S$ (pronounced ``t-c-c'') for a set of receivers $S$.
Both transformations keep the exploded supergraph $G^\#$ the same, and only generate different edge functions.
The solution of the IDE problem
can be mapped back to an IFDS solution.
If the equivalence transformation was used, then this solution is identical to the
solution that would be computed by the IFDS algorithm for the original IFDS problem.
If the correlated-calls transformation was used, then this solution is more precise
because it excludes flow along infeasible paths due to correlated calls.

@@

%\paragraph{Lattice Elements}
In the correlated-calls transformation $\transCC_R$, the 
%Just like in the equivalence transformation $\transEq$, the exploded supergraph for $\transCC$ is the same as in the original IFDS problem. The 
lattice elements are maps from receivers to sets of types:
    $\lcc=\left\{\,m:\ R\to2^T\right\}$,
where $R$ is the set of considered receivers and $T$ is the set of all types.
For each receiver~$r$, the map gives an overapproximation of the possible
runtime types of $r$.
Sets of types are ordered by the superset relation, and this is lifted to maps from receivers
to sets of types, so the bottom element $\botcc$ maps every receiver to the set of all types,
and the top element $\topcc$ maps every receiver to the empty set of types.
During an actual execution, every receiver $r$ points to an object of some
runtime type. Therefore, a data-flow fact
is unreachable along a given path if its corresponding lattice element
maps any receiver to the empty set of types.

@@

A micro-function $f\in\lcc\to \lcc$ defines how the map from receivers to types should
be updated when an instruction is executed.
The micro-function for most kinds of instructions is the identity.
On a call to and return from a specific method $m$ called on receiver $r$,
the micro-function restricts the receiver-to-type map to map $r$ only to
types consistent with the polymorphic dispatch to method $m$.
Finally, when an instruction assigns an object of unknown type to a receiver
$r$, the corresponding micro-function updates the map to map $r$ to the
set of all types. This is made precise by the following definition:
\begin{definition}\label{def:microfn}
Given a previously fixed set $S \subseteq R$ of receivers, the micro-function $\varepsilon_S(e)$
of a supergraph edge $e$ is defined as:
                    %\otodo{define $n_1(e)$ globally somewhere early on, and consider using a more decriptive name like src instead of $n_1$}
  \begin{align}\label{eq:varepsilon}
      &\varepsilon_S(e)= \lambda m\,.\,\\
        &\begin{cases}
            m[r\to m(r)\cap \tau(s,\,f)],
            &\text{\parbox{75mm}{
                    if $e$ is a call-start edge $r.c() \to \textsf{start}_f$ that calls
                procedure $f$ with signature $s$,
                and $r\in S$;}}\\[3ex]
        \parbox{30mm}{\[\begin{array}{l}
                m[r\to m(r)\cap \tau(s,\,f)]\\
            [v_1\to\bot_T] \ldots [v_k\to\bot_T],
        \end{array}
        \]
        }
        &
        \text{\parbox{75mm}{
                if $e$ is an end-return edge
                $\textsf{end}_f \to \textsf{return}_{r.c()}$
                from method $f$ with signature $s$ to the return node corresponding
                to the call $r.c()$,
                $v_1,\,\dots,\,v_k\in S$ are the local variables
                in $f$,
                and $r\in S$;}}\\[5ex]
            m\left[r\to \bot_T\right],
            &\text{\parbox{75mm}{if $e = n_1 \to n_2$ and $n_1$ contains an assignment
                    to $r\in S$;}}\\[2ex]
            m
                &\text{otherwise.}
        \end{cases}
  \end{align}
\end{definition}

@@

\begin{restatable}[Soundness]{lemma}{sound}\label{lem:sound}
  Let $P$ be an IFDS problem and $p=[\startmain,\,\dots,\,n]$ a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  $$
    d\in \backCC\left(\resultIDE(\transCC_R(P))\right)(n)\,.
  $$
\end{restatable}
